---
title: "Second Lesson"
author: "PGaulke"
date: "22 März 2019"
output: pdf_document
---
#Second Lesson

##Task: Building a Book
To change the layout of a book, see output.yml
To change the order of the pdf file, or what ever file: include:

```
bookdown::pdf_book:
  includes:
    in_header: preamble.tex
```

check for special role of bookdown.ywml  
Writing your own webpage, use HTML output form


##Use of the File custom_readthedown

there you can change for example the layout of header etc. (See the Github of Fidalgo for downloading the customreadthedown file)


* Chunks, see Fidalgo Book and its instructions for that.
In brackets there are always these conditions such as eval or echo, and then equal false or true


* to create a new chapter: new r markdown file, put hashtag in the first line to add chapter name, so that it will be displayed in the book as a new chapter (and appears in table of contents)


##Create GitHub account
Download GitHub
Open Global Options and add Git Hub (cread ssh)
Create Repo (same name)
Quick Set Up (see the link)
Add SSH Key



##Calculation Example of R
First put in the chunk as follows, than click on the right arrow to run the chunk
```{r}
2+6
56/6
```
For further Mathoperators, check Fidalgos File


##Data Strucutres
How to store data
Simplest: Vector: just a pile of observations
(for example all the age information of the class)
We have different types of vectors 
* logical: such as false or true
* double: also numeric (age information)
* character: contains word

###How to create a vector:
trop1 as a name, and assigned to the name are the numbers (c is necessary to say: combine these values)
```{r}
trop1 <- c(24,23,23,24,24)
trop1

trop2 <- c("mi","ga","tr")
trop2

trop3 <- c(24,24,"trierweiler")
trop3
typeof(trop3) # to undestand which kind of vector we can use

var <- c(2)
var

tar <- c("2")
tar


```
So now we have two vectors, trop1 which is including age information, and trop2 which includes name information

see differences as definition of character and numeric

**learnings of sec lesson: how to create a book, how to create chapter, how to creater vectors, how to upload on github**

