---
title: "Chapter_01_R_10"
author: "PGaulke"
date: "13 Juli 2019"
output: html_document
---

# Setup

You will need the following software:

The R software itself, RStudio (so to say, the environment where you will work in) and a Latex distribution for creating output files such as pdf files with R graphics.

R Software - https://cran.uni-muenster.de/

FreeVersion of RStudio -  https://www.rstudio.com/products/rstudio/download/#download

Latex distribution - for example: https://www.latex-project.org/get/ (depending on your software)


## Start Your Project

In order to share your work, GitHub is a tool of major importance. In the following, we will explain to you how to set this up. It is totally up to you whether to install it now or later. Please just consider, that you should install it before you start your  project. Otherwise you will face a quite complicated process and limited possibilities to fully enjoy all collaborational features that GitHub provides.

The software you need for this is Git Distribution.

Git distribution - https://git-scm.com/downloads

In parallel of installing the Git distribution, go to https://github.com and create an account.

In a first step, you should activate Git in RStudio. Therefore choose: 'Tools' > 'Global Options' > 'Git/SVN' and click on the button to enable the version control interface. Additionally, you should generate a SHH RSA key which will be needed in a later stage when setting up your repository on GitHub.

Now you can create your project. Click on 'File' > 'New Project' > 'Existing directory' and choose where you want to place your project on your computer. Be aware, that you really know where you place it as you will need the directory in a later stage.

Now it is time to prepare for the marriage of your GitHub account and your project. Go to GitHub and create a new repository and name it exactly the same way as you named your R project. The naming has to be identical.
Next, got to your settings in GitHub and choose 'SHH and GPG Keys' and click on 'New SHH key'. Go back to RStudio and copy the SHH key that has been created in the first step, then paste it into your GitHub account.

Now everything is set up to create the connection. In order to do so, go in RStudio to 'Tools' > 'Project Options' > 'Git/SVN'. Select 'Git' in the Version control system field. After that, go again to 'Tools' > 'Terminal' > 'New Terminal'. Now next to the console a terminal should appear.

As you may already read when you finished your the creation of your repository, here you should type in the following commands:


```{r, eval=FALSE}
git init


git remote add origin https://github.com/YOURNAME/YOURREPOSITORY.git
git push -u origin master


 
```

Obvisously you should adopt the origin link with your own names. 

Now restart RStudio and enjoy that you just completed to connect your project with GitHub. In the upper right corner of RStudio you should now see a Git button (right next to environment/history/connections). To put your files on GitHub, you can now easily commit and then push all files you want to share.

## Work Collaboratively

If you are interested to work simoultanesly with another person on one project, you can create a team in GitHub. However, this is not done by a few clicks.

At first, you have to decide who the owner of the project should be. The role will not have too much influence later on, but it defines the set up for each participant.

The project owner needs to create an organization on Github. Within this organization a new repository should be created, which again should have exactly the same name as the project in R Studio. Herefore, you can follow the steps described above.
If you have done that, you should create a team in the organization and add the further participants to the team. Be aware that you should assign the repository to these members and provide the members with respective rights. 

As the team member who should have received an email at this point. After you confirmed the participation, go onto the repository and click on the button 'Clone or download'. Look for the https adress, copy it and go now into a simple RStudio session (not a project). Open a new Terminal and type in cd with the path where you want to save the project.
Now typ 'git clone' and paste the the copied link from GitHub.

```{r, eval=FALSE}
cd Dropbox/Master/2Sem
git clone https://github.com/ORGANIZATION/repo.git 
```

Now you should finally be able to push and pull the all files and start your collaborative R project.