---
title: "05_Fifith_Lesson"
author: "PGaulke"
date: "10 April 2019"
output: pdf_document
---
#Fifth Lesson

##Sub-Setting
####3.1 Excercise on Subsetting
Subsetting with Indices

```{r}
#a=(a)
alphabet <-c(letters)

alphabet[1]#a

alphabet[1:3]#b

alphabet[c(-2,-3)]#c

#d dont know

alphabet[c(TRUE,FALSE,FALSE)]

```

####3.2 Excercise on Subsetting
Subsetting with Conditions

```{r}
set.seed(30)#i
set.seed

data <- c(runif(50, min = 80, max = 120))#ii
data

#iii
length(data[data<90])

#iv
#dont know

#v
which(data%%2==0)

#vi
data[data>(mean(data))]

#vii
#dont know

```

####3.3 Excercise on Subsetting
Subsetting data frames

```{r}
#i
A <-data.frame(iris)

#ii
max(iris$Sepal.Length)


#iii

#iv
data.frame(iris$Sepal.Length,iris$Sepal.Width)
#v

#vi

#vii



```

##Sub-Dialecct of R - Tidyverse

Download packages tidyverse, tiddle

Than: library(tidyverse) --> run this 

###Use of tibble

With making a dataset a tibble (see chunk below) it is possible to define the Spalten. For example it is easily possible to create different columns by using mathematically mixed informations

Example: Column 1 = x, Column 2 = y, Column 3 = xy




```{r}
library(tidyverse)
t_iris <- tibble(iris)
class(t_iris)


```



###magrittr

How to pipe= %>%

```{r}

#simple

library(magrittr)
mydata <- c(2:10,20,42)
mean(mydata)

mydata%>%
  mean()

#advanced

round(sqrt(mean(mydata)),2)

mydata %>%
  mean() %>%
  sqrt() %>%
  round(2)


##using a placeholder

3 %>%
  round(17.2342)


```


Transformations on a data frame

df_after_f <-f(df)
df_after_g <-g(df_after_f)
df_after_h <-g(df_after_g)

with piping it is

df %>%
  f %>%
  g %>%
  h
  

Own example
```{r}

library(magrittr)
library(tidyverse)


data (iris)
t_iris <- as_tibble(iris)

t_iris2 <- t_iris[t_iris$Sepal.Length>4.5 & t_iris$Sepal.Length<5.5,]
t_iris2

av.sw<-mean(t_iris$Sepal.Length)
av.sw

iris %>%
  as_tibble() %>%
  .[.$Sepal.Length==5,] %>%
  {mean(.$Sepal.Width)}



```

###How to Tidy Up something: Install tidyr

how to gather
```{r}

lottery <- tibble (person= c("Andre","Beth","Charles"),
  mon= c(2,5,1),
  tue= c(3,7,5),
  wed= c(10,3,4),
  thu= c(1,2,7)
  )

lottery



#tidy it up

tidy_lottery <- gather (lottery,
                        key= "day",
                        value= "gains",
                        mon:thu
                        )
tidy_lottery



#another example
bisou <- tribble(
  ~name, ~`2016`, ~`2017`, ~`2018`,
  "gau", 23,45,12,
  "tri", 8,14,30,
  "han", 12,43,8
)
  
bisou

tidy_bisou <- gather(bisou,
                     key="Year",
                     value="Number",
                     `2016`, `2017`, `2018`)

tidy_bisou

```


how to spread
```{r}
people<- tibble (name=rep(c("Andre", "Beth", "Charles"),2),
                 info=rep(c("eyes","age"),3),
                 measure= c("Blue",23,"brown",31,"brown",19)
)
                 
                 
tidy_people<- spread(people, key=info, value=measure)
tidy_people

people2<- tibble (name=c(rep(c("Andre", "Beth", "Charles"),2),"David"),
                 info=c(rep(c("eyes","age"),3),
                 "height"),
                 measure= c("Blue",23,"brown",31,"brown",19, 175)
)
                 
                 
tidy_people2<- spread(people2, key=info, value=measure)
tidy_people2

```

<<<<<<< HEAD
#Unite 
#Exercises on sub-setting

```{r}
letters 
v1 <- letters [1]

v2 <- letters [c(1,2,3)]
v2

v3 <- letters[-c(2,3)]


v4 <- letters[-c(24, 25, 26)] # bad one

v4 <- letters[-c(length(letters)-2, length(letters)-1, length(letters))]
lv <- length (letters)



v5 <- letters[seq(from=1, to=26, by 3)]
v5
```

```{r}
data (iris)
max(Sepal.Lenght) # not found bc. it is in the data frame of iris, tell them, it should look in the data frame
```

```{r}
data (iris)
max(iris$Sepal.Length)
```

# iii

```{r}
iris2 <- iris [, c("Sepal.Length", "Sepal.Width")]
```

=======
how to separate
```{r}
clients <- ("Andre, Porto")


```


VERY IMPORTANT: EVERYTIME WE WANT TO USE TIBBLE OR PROCESS TIBBLE (OR ANY OTHER PART OF TIDYVERSE) WE NEED TO PUT IN THE CONSOLE: library(tidyverse) (BECAUSE TIDYVERSE IS A SUBLANGUAGE OF R/WE DIDNT DO ANY OTHER LANGUAGE SO FAR, JUST BASIC R)
>>>>>>> b5b9a9d2087ef2869d0701f2b10cd730c6eb23f3


