---
title: "05_Fifith_Lesson"
author: "PGaulke"
date: "10 April 2019"
output: pdf_document
---
#Fifth Lesson

##Sub-Setting
####3.1 Excercise on Subsetting
Subsetting with Indices

```{r}
#a=(a)
alphabet <-c(letters)

alphabet[1]#a

alphabet[1:3]#b

alphabet[c(-2,-3)]#c

#d dont know

alphabet[c(TRUE,FALSE,FALSE)]

```

####3.2 Excercise on Subsetting
Subsetting with Conditions

```{r}
set.seed(30)#i
set.seed

data <- c(runif(50, min = 80, max = 120))#ii
data

#iii
length(data[data<90])

#iv
#dont know

#v
which(data%%2==0)

#vi
data[data>(mean(data))]

#vii
#dont know

```

####3.3 Excercise on Subsetting
Subsetting data frames

```{r}
#i
A <-data.frame(iris)

#ii
max(iris$Sepal.Length)


#iii

#iv
data.frame(iris$Sepal.Length,iris$Sepal.Width)
#v

#vi

#vii



```

##Sub-Dialecct of R - Tidyverse

Download packages tidyverse, tiddle

Than: library(tidyverse) --> run this 

###Use of tibble

With making a dataset a tibble (see chunk below) it is possible to define the Spalten. For example it is easily possible to create different columns by using mathematically mixed informations

Example: Column 1 = x, Column 2 = y, Column 3 = xy




```{r}
library(tidyverse)
t_iris <- tibble(iris)
class(t_iris)


```



###magrittr

How to pipe= %>%

```{r}

#simple

library(magrittr)
mydata <- c(2:10,20,42)
mean(mydata)

mydata%>%
  mean()

#advanced

round(sqrt(mean(mydata)),2)

mydata %>%
  mean() %>%
  sqrt() %>%
  round(2)


##using a placeholder

3 %>%
  round(17.2342)


```


Transformations on a data frame

df_after_f <-f(df)
df_after_g <-g(df_after_f)
df_after_h <-g(df_after_g)

with piping it is

df %>%
  f %>%
  g %>%
  h
  

Own example
```{r}

library(magrittr)
library(tidyverse)


data (iris)
t_iris <- as_tibble(iris)

t_iris2 <- t_iris[t_iris$Sepal.Length>4.5 & t_iris$Sepal.Length<5.5,]
t_iris2

av.sw<-mean(t_iris$Sepal.Length)
av.sw

iris %>%
  as_tibble() %>%
  .[.$Sepal.Length==5,] %>%
  {mean(.$Sepal.Width)}



```

###How to Tidy Up something: Install tidyr


```{r}

lottery <- tibble(person= c("Andre","Beth","Charles"),
  mon= c(2,5,1),
  tue= c(3,7,5),
  wed= c(10,3,4),
  thu= c(1,2,7)
  )

lottery



#tidy it up

tidy_lottery <- gather (lottery,
                        key= "day",
                        value= "gains",
                        mon:thu
                        )
tidy_lottery



#another example
bisou <- tribble(
  ~name, ~y
  
  
  
)

```



