---
title: "06_Sixth_Lesson"
author: "PGaulke"
date: "24 April 2019"
output: pdf_document
---

#Sixth Lesson

##Tidying
####5.1 Excercise on Tidying
Transform 'data' into 'data_tidy' by using using `gather` then `separate` then `spread`.

```{r}
data <- tribble(~famid, ~named, ~incd, ~namem, ~incm, 
                1, "Bill", 30000, "Bess", 15000,
                2, "Art", 22000, "Amy", 18000,
                3, "Paul", 25000, "Pat", 50000
                ) 
data


data_tidy <- gather(data, 
                    key= "famid",
                    value= "incd"
                    )

data_tidy
```



##Start of the lesson
###dplyr
easier and fast way to manipulate data
this is another part of the sublanguage tidyverse, so again put in libraby(tidyverse) in the console
(and of course download dplyr)

First verbes: SELECT and MUTATE that can be applied on colunns

```{r}
starwars %>%
  select(name, species)


# compare with base R
starwars [, c("name","species")]

starwars %>%
  select(
    name, 
    matches("color")
    )


```
### Helper arguments

These are arguments to help `select` the variables/columns. Here are a few of them (see `?select` for a complete list):

- `starts_with("x")`
- `ends_with("x")`
- `contains("x")`
- `matches("x")`


```{r}
starwars %>%
  select(name, gender, films)
```

Now switching from SELECT to MUTATE

```{r}
starwars %>%
  mutate(bmi = mass / (height/100)^2)

df_b <- starwars %>%
  mutate(
    bmi = mass/ (height/100)^2,
    sr.bmi =sqrt(bmi)
  )

df_b

df_a <- starwars %>%
  mutate(g_b = sample(c("G","B"), 87, replace = TRUE)
  )
df_b
```

Now switching to FILTER (for rows now, and not for columsn as for SELECT and MUTATE)

```{r}
starwars %>%
  filter (
    species !="human" &
    eye_color == "red"
  )

starwars


# in base r
starwars[starwars$species!="Human" &
           starwars$eye_color=="red",]

```

now switch to ARRANGE (also rows)
```{r}
starwars %>%
  arrange(
    height,
    desc(mass)
  )

starwars %>%
  arrange(
    desc(name)
  )

starwars %>%
  arrange(
    name
  )

```

now switch to SUMMARISE
```{r}
#creates a new df by collpasing values
starwars %>%
  summarise(
    min_h = min(height, na.rm = TRUE),
    av_m=mean(mass, na.rm =TRUE)
  )

#site comment: to remove not available data, na.rm =TRUE
```
```{r}
#combine the verbs
starwars %>%
  select(species, mass) %>%
  filter(species == "Human") %>%
  summarise(av_m=mean(mass, na.rm = TRUE))

#in regular R language
summarise(filter(starwars, species =="Human"),
          av_m=mean(mass, na.rm=TRUE))


```
