---
title: "Second Lesson"
author: "PGaulke"
date: "22 März 2019"
output: pdf_document
---
#Second Lesson

* Task: Building a Book
* To change the layout of a book, see output.yml
To change the order of the pdf file, or what ever file: include:

```
bookdown::pdf_book:
  includes:
    in_header: preamble.tex
```

* check for special role of bookdown.ywml  
    
    
* Writing your own webpage, use HTML output form
use the file: custom_readthedown
there you can change for example the layout of header etc. (See the Github of Fidalgo for downloading the customreadthedown file)


* Chunks, see Fidalgo Book and its instructions for that.
In brackets there are always these conditions such as eval or echo, and then eqaul false or true


* to create a new chapter: new r markdown file, put hashtag in the first line to add chapter name, so that it will be displayed in the book as a new chapter (and appears in table of contents)

* Create GitHub account
Download GitHub
Open Global Options and add Git Hub (cread ssh)
Create Repo (same name)
Quick Set Up (see the link)
Add SSH Key

* How to Use are, for example as a single calculator
First put in the chunk as follows, than click on the right arrow to run the chunk
```{r}
2+6
56/6
```
For further Mathoperators, check Fidalgos File

*next Chapter: Data structures
How to store data
Simplest: Vector: just a pile of observations
(for example all the age information of the class)
We have different types of vectors 
+ logical: such as false or true
+ double: also numeric (age information)
+ character: contains words

* how to create a vector:
trop1 as a name, and assigned to the name are the numbers (c is necessary to say: combine these values)
```{r}
trop1 <- c(24,23,23,24,24)
trop1

trop2 <- c("mi","ga","tr")
trop2

trop3 <- c(24,24,"trierweiler")
trop3
typeof(trop3)

var <- c(2)
var

tar <- c("2")
tar

var+30
tar+30

```
So now we have two vectors, trop1 which is including age information, and trop2 which includes name information

to show which kind of vector we can use typeof


see differences as definition of character and double

* learnings of sec lesson: how to create a book, how to create chapter, how to creater vectors, how to upload on github

