---
title: "09_Tenth_Lesson"
author: "PGaulke"
date: "30 Mai 2019"
output: html_document
---

#Tenth Lesson


```{r}
advertising <- read.csv("data/Advertising.csv")

advertising

```

##Dataset


see the Graph Credit Card Default. This is what we want to reproduce now.

```{r}
require(ISLR)
require(tidyverse)
#to load the packages that we need


as_tibble(Default)
#to make irt

faster

str(Default)
Default
# to see the structure

```

```{r}
#now drawing the graph

#basic way:
#plot(Default$balance,Default$income)

#now with colour
#plot(Default$balance,Default$income, col = Default$default)

#to decrease the size of the data set, get a sample out of the whole set
#show.index <- sample(1:length(Default$income),1000)
#or
show.index <- sample(1:nrow(Default),1000)
show.index

#final
plot(Default$balance[show.index], Default$income[show.index], col=Default$default[show.index])



```


##now the boxplo{r}

```{r}
boxplot(Default$balance ~ Default$default, col = "red")



```

##measure the accuracy on this non-numeric

Instead of RSME (just numeric)


```{r}

table(Default$default)/nrow(Default)*100

#reason why this is pracitcally a bad way to predict, (drawed matrice in the paper that we received)

#Sensitivity
#is better to, to check any variable (default or not default) seperately (other wise, all "save" guesses will increase unrealisticly the accuracy result)

```

##Why not linear regression

```{r}
#how we create the linear regression
#modz <- lm(default ~ balance, data = Default)
#will produce error, therefore change default to a numeric set

modz <- lm(as.numeric(default) ~ balance, data =Default)
summary(modz)

#now default is not yes or now, but 1 or 2, therefore numeric

#repetition: fitted values shows the data of a graph as a prediction: the streight line which is created
#now we draw

plot(Default$balance, Default$default)
lines(Default$balance, modz$fitted.values, col = "blue")
#modz$fitted.values

```




```{r}
levels(Default$default)
predicted.class <- ifelse(modz$fitted.values < 1.19,"No","Yes")

#predicted.class is a vector, ifelse is a function: if the condition is met go for yes, if not, go for no
#herefore we use the value of fitted values (which is basically the predicted value)

#to now use accuracy or sensitvity model, make a table 
table(predicted.class, Default$default)

table(factor(predicted.class, levels=c("No", "Yes") ),Default$default)

#thats all whats on the sleft side of the slide (printed out material, linear versus logistic regression)



```
###Logistic Regression

```{r}
#now we start to draw the right line (logistic regression)
#we are orientieren an cumulative distribution

#function that we us is glm (generealized linear model)

#we want to estiamte the default with balance

mod.pressure <-glm (default ~ balance, data = Default, family = binomial)
mod.pressure


#summary(mod.pressure)

#now draw it

#first plot, than check the accuracy


#we dont hav fitted values, as before, therefore we need to create fitted values first
#first create a new data frame, where the predict function gets the new data to check the actual prediction

bal.pre <- data.frame(balance= seq(0,max(Default$balance), length.out = 100))

#see the sequence formula to take 100 values from the min to maximum, with regular distance to each other
fit_glm <- predict(mod.pressure, newdata = bal.pre, type ="response")
fit_glm

#plot (bal.pre$balance, fit_glm, type ="l", col="blue")

#now points and lines
#plot(Default$balance, Default$default)
#lines(bal.pre$balance, fit_glm+1, col ="blue")

#now check accuracy

predicted.class.glm <- ifelse(mod.pressure$fitted.values < 0.5 ,"No","Yes")
table(predicted.class.glm, Default$default)
```

###now with more variables than just balance, so also income and student
```{r}
mod.pressure2 <-glm (default ~ balance + income + student, data = Default, family = binomial)

summary(mod.pressure2)

predicted.class.glm2 <- ifelse(mod.pressure2$fitted.values < 0.5 ,"No","Yes")
table(predicted.class.glm2, Default$default)
```

now try how to get the graph of the slide confounding (students and probability of credit default)


```{r}
mod.pressure3 <-glm (default ~ balance + student, data = Default, family = binomial)

summary(mod.pressure3)

#first step: create a data.frame for whcih we want to predict
var.preYes <- data.frame(balance= seq(0,max(Default$balance), length.out = 100), student= "Yes")
var.preNo <- data.frame(balance= seq(0,max(Default$balance), length.out = 100), student= "No")

fit_glm3y <- predict(mod.pressure3, newdata = var.preYes, type ="response")
fit_glm3y

fit_glm3n <- predict(mod.pressure3, newdata = var.preNo, type ="response")
fit_glm3n


plot(Default$balance, Default$default)
lines(var.preYes$balance, fit_glm3y + 1, col="red")
lines(var.preNo$balance, fit_glm3n + 1, col="blue")


```