---
title: "09_Third_Lesson"
author: "PGaulke"
date: "29 März 2019"
output: pdf_document
---

#Third Lesson
First Topic: Data Structure

verschiedene Data Strucutre (arbeiten mit Vektoren)

```{r}
# populate a matrix with elements in of the vector, and give the dimension

M <- matrix(c(2,3,4,5,6,7),nrow=3)

M
```


Das obige ist um eine Matrix bzw. Tabelle zu erstellen




```{r}
school<-factor(c("PRimary","Secondary","Tertiary"), ordered=TRUE)
mylist <-list (numbers=c(1:60),
              somenames=c("jim", "jules"),
              results=c(T,F,F,T),
              school=school)

mylist
```

Das obige ist um eine Liste zu generieren.


```{r}
df <- data.frame(let=LETTERS[1:7],
                 num1=10:16,
                 num2=floor(rnorm(7,100,10)))
df
names(df)
```
```{r}
students <- data.frame(name=c("Corinna", "Philipp", "Freddy", "Moritz", "Niklas", "Laurin", "Forster", "Fidalgo"),
                       age=c(23, 23, 24, 22, 23, 24, 30, 30),
                       likeR=c(T,F,T,F,T,F,T,F))
                    
students
length(students)
```


How are Functions by vectors?
Element by Element Evaluation

```{r}
# create two numeric vetor of same length
visits1 <- c(12,2,45,75,65,11,2)
visits2 <- c(23,4,5,78,12,0,200)
length(visits1)
length(visits1) == length(visits2)

total <- visits1 + visits2
# their sum is the vecotr of same length, with the element by element sums

length(total)
total
```

above: element by element evaluation, weil jedes element (erster Zeitraum bei visit1 plus erster Zeitraum bei visit2) addiert wird

```{r}
total.p2 <- total^2
total.p2

(trop <- total - 2)

```

total p2 is just the name and totally replacable. No reason for the name, everything after is not replacable for this calc




Another rule: Recyling

```{r}
l3 <- c(12,34,50)
l2 <- c(10,3)
tt <- l3 + 5
tt

ltotal <- l3+l2 # recycling!
ltotal

```

recycling happens than, the red is not really an error message but it makes us aware that recycling happens (10+50)


Attention: SOMETIMES RECYCLING AND ELEMENT BY ELEMENT EVALUATION HAPPENS




Next Topic: To get help
Type in into the console: ?mean to get the description in the help window

Then this appears in the top:
mean {base}	R Documentation
Arithmetic Mean

{base} shows the package

```{r}
ages <- c(28,33,39,56,34,45,27,40,NA)
ages
mean(ages, na.rm = T, trim = 30,40)


```

na.rm = t used to ignore the NA in the calc. Otherwise it would be an error message. na.rm = T stands for: ignoring NA. (By saying =false than we have the default (same as for writing nothing, so not na.rm at all))


#Excercise 2 on simple vectors creation

2.1creation of simple vectors

## Creation of simple vectors

Create the following simple vectors.

  - \(a = (1, 2, 3, ..., 15)\),
  - \(b = (15, 14, 13, ..., 0, -1, -2)\),
  - \(c = (15, 14, 13, ..., 10, 9, -9, -10, ..., -13, -14, -15)\)
```{r}
#2.1 create vector with numbers
a <- c(1:15)
b <- c(15:-2)
c <- c(15:9,-9:-15)

a

b

c
```

2.2 ## Empty vector

Create an empty vector with 30 numeric values. Check its type and its length.   
(Hint: use `numeric()`.)


```{r}
# 2.2 empty vector
empty <- numeric(c(30))


empty

length(empty)
typeof(empty)

# or

empty2<- numeric(length=30)

empty2
```

2.3 ## Character vector

Create the vector `participants` with the names of the participants in the class.

```{r}
# 2.3 List all participants
participants <- c("corinna", "philipp", "laurin", "freddy", "moritz", "niklas")

participants
```
 
 
 2.4
Consider the following scores at a test.
\[(60, 84, 65,67,75,72,80,85,63,82,70,75)\]
Calculate the mean of the sample.  
Recall the formula for the z-score of an observation:
\[z_i=\frac{x_i-\mu}{\sigma}\]
Create a vector with the z-score of each grade. 

```{r}
# 2.4 simple manipulation 
scores <- c(60, 84, 65, 67, 75, 72, 80, 85, 63, 82, 70, 75)

mean(scores)
sd(scores)
z.scores <- (scores-mean(scores))/sd(scores)
z.scores


```

2.5 ## Simple calculations

Calculate the following expressions, where \(n\) is the number of elements in the series.
\[\frac{1}{n}\sum_{i=5}^{55} (i^2 - 3i)\]
\[\frac{1}{n}\sum_{i=1}^{10} \left(\frac{3^i}{3i} + \sqrt{3i}\right)\]

```{r}

i <- c(5:55)

aa<- i^2-3*i
length(aa)

sum(aa)/length(aa)

#better

sum (sum (i^2-3*i))/length(i)


```


#Subsetting

Subsetting with [ and ] - square brackets

to get specific part of the set


```{r}
va <- c(13.1, -15.2, 0.3, 2.4, 10.5, -3.6, 9.7)
va[1] # first element
va[c(3:5)] # element 3 to 5
va[-1] # all elements, but without element 1
va[c(T,T,T,F,F,T,F)] #just elements 1,2,3 and 6
va [c(T,F)] #notice the recycling
va[c(-2,-7)] #all but second and seventh




```