---
title: "10_Fourth_Lesson"
author: "PGaulke"
date: "5 April 2019"
output: pdf_document
---

#Fourth Lesson


2.7
For the following questions, you may want to check the help on the function seq.
Create the following vectors:
• a = (1, 2, 3, ..., 15),
• b = (15, 14, 13, ..., 0,−1,−2),
• c = (1, 1.2, 1.4, ..., 8).
Create the following vectors, potentially based on auxiliary vectors.
• d = (0.531.31, 0.561.34, ..., 0.5361.334),
• e = (5, 52, 2 , 53, 3 , ..., 525
25 ).
Use the function all.equal to check that the last elements of each vector are indeed 1.0889154 × 10−7 and
1.1920929 × 1016, respectively.


```{r}
a <- c(1:15)
b <- c(15:-2)

x0=1
c <- seq(x0,8,0.2)
c

xtfrm


```


2.8
 Use paste in vector creation
For the following questions, you may want to check the help on the function paste.
Create the following vectors:
• a = (City 1, City 2, . . . , City 6),
• b = (n1, n2, . . . , n6)
• b = (Gender=M, Gender=F, Gender=M, Gender=F, Gender=M, Gender=F)
3


```{r}




```


#What happens in the lesson:

Differences between matrice and normal vectors. How we "access" or "subset" a matrice or vector: Vector just a list, Matrice has different dimension. Therefore, we need: mat.(row, column) to display what we want.
Shorter writing style for getting all values of a column: matrice$column
Than you can subset this matrice.

How to get a subset out of a matrice. (BTW r condition has no logic, it is more a name)

```{r condition}
mtcars
mtcars[mtcars$mpg<20,] # all cars with mpg smaller 20


```

Further on vectors, put in conditions to display something like a statement

```{r}

gains <- c(10,3,-5,-4,12,4)
sum(gains)

# change the element -4 to -40 with a formula
# subset gains and assign new values
gains[gains==-4]<- -40
gains

if (sum(gains) >0) {
  print("Congratulations, you are winning")
} else  {
  print ("You are not winning")
}
# else statement for display something else in case the if condition is false

```


Less trivial ways of writing condition
  
```{r}
gains <- c(10,3,-5,0,-4,12,4)

if (is.numeric(gains)) {
  print ("Seems like it is a numeric vector...")
}

# here is-numeric (gains) evaluates to true hece the code is exceuted.

!is.data.frame(gains)

class(gains)

```


##Structures of simple function
How to write functions


pow_two is just a random name here
```{r}
pow_two <- function (a){
  a^2
}
pow_two(12)
```

you can also here apply element by element evaluation



}
