---
title: "13_Solution_MLR"
date: "26 Juli 2019"
output: html_document
---


```{r}
require(tidyverse)
require(ISLR)
require(magrittr)


load("C:/Users/admin/Dropbox/Master/2. Semester/Data Science/MyBook/project_data.Rdata")
summary(train.data)
summary(test.data)

# 11 variables for frequency of seven plants
# task: The test.data has the same structure but does not contain the frequencies for each of the 7 plants. Your
# goal is precisely to estimate them for the 140 observations

# variables:character to numeric
# for train.data
train.data$season = as.numeric(c("spring" = "1", "summer" = "2", "autumn" = "3", "winter" = "4")[train.data$season])

train.data$size = as.numeric(c("small" = "1", "medium" = "2", "large" = "3")[train.data$size])

train.data$speed = as.numeric(c("low" = "1", "medium" = "2", "high" = "3")[train.data$speed])


# for test.data

test.data$season = as.numeric(c("spring" = "1", "summer" = "2", "autumn" = "3", "winter" = "4")[test.data$season])

test.data$size = as.numeric(c("small" = "1", "medium" = "2", "large" = "3")[test.data$size])

test.data$speed = as.numeric(c("low" = "1", "medium" = "2", "high" = "3")[test.data$speed])


# delete na
#is.na(train.data)
#na.omit(train.data)

# replace missing value with the mean (and avoid a loop!)

train.data$mxPH[is.na(train.data$mxPH)] <- round(mean(train.data$mxPH, na.rm = TRUE))
train.data$mnO2[is.na(train.data$mnO2)] <- round(mean(train.data$mnO2, na.rm = TRUE))
train.data$Cl[is.na(train.data$Cl)] <- round(mean(train.data$Cl, na.rm = TRUE))
train.data$NO3[is.na(train.data$NO3)] <- round(mean(train.data$NO3, na.rm = TRUE))
train.data$NH4[is.na(train.data$NH4)] <- round(mean(train.data$NH4, na.rm = TRUE))
train.data$oPO4[is.na(train.data$oPO4)] <- round(mean(train.data$oPO4, na.rm = TRUE))
train.data$PO4[is.na(train.data$PO4)] <- round(mean(train.data$PO4, na.rm = TRUE))


# to to the same with the test.data
test.data$mxPH[is.na(test.data$mxPH)] <- round(mean(test.data$mxPH, na.rm = TRUE))
test.data$mnO2[is.na(test.data$mnO2)] <- round(mean(test.data$mnO2, na.rm = TRUE))
test.data$Cl[is.na(test.data$Cl)] <- round(mean(test.data$Cl, na.rm = TRUE))
test.data$NO3[is.na(test.data$NO3)] <- round(mean(test.data$NO3, na.rm = TRUE))
test.data$NH4[is.na(test.data$NH4)] <- round(mean(test.data$NH4, na.rm = TRUE))
test.data$oPO4[is.na(test.data$oPO4)] <- round(mean(test.data$oPO4, na.rm = TRUE))
test.data$PO4[is.na(test.data$PO4)] <- round(mean(test.data$PO4, na.rm = TRUE))

```

```{r, include = FALSE}
#create test data
set.seed(200)
num_obs = nrow(train.data)

train.index = sample(num_obs, size = trunc(0.50 * num_obs))
newtrain.data = train.data[train_index, ]
traintest.data = train.data[-train_index, ]


# delete na
is.na(newtrain.data)
na.omit(newtrain.data)

is.na(traintest.data)
na.omit(traintest.data)


newtrain.data
traintest.data
```

## Develop the mlr model for a1 - Linear Regression

Determining the signifance level wanted: <0,05
```{r}

# Linear regression on multiple linear regression

mlrmod <- lm(a1 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=train.data)
summary(mlrmod)


```

The p-value of the F-Statistic is 7.936e-11, therefore it is highly signifcant and at least one of the variables is signifcantly related to the outcome response. Clarifying which ones are significant:



```{r}
summary(mlrmod)$coefficient
```


Deleting the highest P values
season, mxPH, oPO4

```{r}
# the new one

mlrmod2 <- lm(a1 ~   size + speed + mnO2 + Cl + NO3 + NH4 + PO4 + Chla, data=train.data)
summary(mlrmod2)

# delete speed, mn02, Cl, Chla, NH4

mlrmod2 <- lm(a1 ~   size + NO3 + PO4, data=train.data)
summary(mlrmod2)

# predict
a1pred <- predict(mlrmod2, test.data)
a1pred


test.data$a1<-a1pred


# all negative values to zero
test.data$a1[test.data$a1<0] <- 0
test.data$a1


summary(test.data)

```

__________________________________________________________________________________________________________


##  Develop the mlr model for a1 - Polynomials

```{r}

# Linear regression on multiple linear regression

polymod <- glm(a2 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=train.data)
summary(polymod)

```

Deleting variables with lowest significance.

mnO2, Cl, season, NO3, PO4, OPO4, NH4, size

```{r}

# Linear regression on multiple linear regression

polymod2 <- lm(a2 ~ speed + mxPH + Chla, data=train.data)
summary(polymod2)

# use of poly

polymod3 <- lm(a2 ~ poly(Chla,3)+ poly(speed,2) + poly(mxPH,2), data=train.data )
summary(polymod3)

a2pred <- predict(polymod3,test.data)
summary(a2pred)

```

Try to put Chla (only) poly in a graph
```{r}

fita2 <- lm(a2 ~ poly(Chla,10), data=train.data )
summary(fita2)

fita3 <- lm(a2 ~ poly(Chla,5), data=train.data )
summary(fita3)

fita4 <- lm(a2 ~ Chla, data=train.data )
summary(fita4)

# poly just on Chla

train.data <- train.data %>%
	mutate(fit1 = predict(fita2),
	fit2 = predict(fita3),
	fit5 = predict(fita4))

cols <- c("Deg.3", "Deg.2", "Deg.1")
train.data %>% 
	ggplot(aes(x=a2, y=Chla)) +
	geom_point() +
	geom_line(aes(y=fit1, color="Deg.3"), size =1) +
	geom_line(aes(y=fit2, color="Deg.2"), size =1) +
	geom_line(aes(y=fit5, color="Deg.1"), size =1) +
	theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```

```{r}
a1pred <- predict(lmod, traintest.data)

summary(a1pred)

actuals_pred <- data.frame(cbind(actuals=newtrain.data$a1, predicteds=a1pred))

RSS <- sum((traintest.data$a1 -  predict(a1pred, traintest.data))^2)
RSS


correlation_accuracy<- cor(actuals_pred)
head(actuals_pred)

```





## Developing a model for a3 - Logistic regression


```{r}
logmod <- as.formula("a3 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla")

logit <- glm(logmod, data = train.data) # Ausgabe der Schätzergebnisse 
summary(logit)

logitMinAIC <- step(logit) 
summary(logitMinAIC)

a3pred <- predict(logitMinAIC,test.data)
a3pred

train.data$a3
```

- speed  1   8862.8 1333.8
- NH4    1   8891.7 1334.5
- mnO2   1   9506.9 1347.9

However, this method is not suitable for this case as K>2 respectively the response is any numeric.

## Developing a model for a4 - Tree

```{r}
require(rpart)

treemod <- rpart(a4 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = train.data)

minofcp <- treemod$cptable[which.min(treemod$cptable[,"xerror"]),"CP"]

treemod2 <- prune(treemod, cp = minofcp)
summary(treemod2)

plot(treemod2)
text(treemod2, pretty = 0)

a4pred <- predict(treemod2, test.data)
a4pred
summary(a4pred)

```

## Developing a model for a5 - RandomForest (Boosting? Bagging? RandomForest?Resampling (here bootstrapping?)


```{r}

require(randomForest)
forestmod <- randomForest(a5 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = train.data, na.action = na.omit)

forestmod



```
