---
title: "12b_Solution_a2"
author: "CTrierweiler,PGaulke"
date: "29 Juli 2019"
output: html_document
---

# Develop the Model for a2 

Methods that will be tested:

- Multiple Linear Regression
- Linear regression with polynomials
- Log Linear Regression
- Trees
- RandomForest

Logistic regression is not applicable on this data set as the response is not binary.

## Multiple Linear Regression

Determining the signifance level wanted: <0,05
```{r}
# Test it

a2.testmlrmod <- lm(a2 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a2.testmlrmod)

# Chla with highest significance

a2.testmlrmod$coefficients["Chla"]

a2.test.pred.mlr <- (predict(a2.testmlrmod,traintest.data))
 
a2.RMSE.testmlr <- sqrt(mean((a2.test.pred.mlr-traintest.data$a2)^2))
a2.RMSE.testmlr

a2.MAE.testmlr <- mean(abs(a2.test.pred.mlr-traintest.data$a2))
a2.MAE.testmlr

# Reducing the variables
a2.testmlrmod2 <- lm(a2 ~ speed + mxPH + Chla, data=newtrain.data)

summary(a2.testmlrmod2)
a2.testmlrmod2$coefficients["Chla"]

a2.test.pred.mlr2 <- (predict(a2.testmlrmod2,traintest.data))
 
a2.RMSE.testmlr2 <- sqrt(mean((a2.test.pred.mlr2-traintest.data$a2)^2))
a2.RMSE.testmlr2

a2.MAE.testmlr2 <- mean(abs(a2.test.pred.mlr2-traintest.data$a2))
a2.MAE.testmlr2

# Reducing the variables to only one variable (SLR)

a2.testmlrmod3 <- lm(a2 ~ Chla, data=newtrain.data)

summary(a2.testmlrmod3)

a2.test.pred.mlr3 <- (predict(a2.testmlrmod3,traintest.data))
 
a2.RMSE.testmlr3 <- sqrt(mean((a2.test.pred.mlr3-traintest.data$a2)^2))
a2.RMSE.testmlr3

a2.MAE.testmlr3 <- mean(abs(a2.test.pred.mlr3-traintest.data$a2))
a2.MAE.testmlr3

# MLR2 better, optimize
a2.testmlrmod4 <- lm(a2 ~ mxPH + Chla, data=newtrain.data)

a2.test.pred.mlr4 <- (predict(a2.testmlrmod4,traintest.data))
 
a2.RMSE.testmlr4 <- sqrt(mean((a2.test.pred.mlr4-traintest.data$a2)^2))
a2.RMSE.testmlr4

a2.MAE.testmlr4 <- mean(abs(a2.test.pred.mlr4-traintest.data$a2))
a2.MAE.testmlr4

# a2-testmlrmod4 as best choice

```

Trying k-fold cross validation

```{r, error=TRUE}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a2.crossmodelmlr <- train(a2 ~ speed + Chla, data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a2.crossmodelmlr)
```


## Polynomials

Adding polynomials

```{r}

# test it from mlr4

a2.testpoly <- glm(a2 ~ speed + Chla, data=newtrain.data)
summary(a2.testpoly)
# use of poly

a2.testpolymod1 <- glm(a2 ~ poly(as.numeric(speed),2) + poly(Chla,2), data=newtrain.data)
summary(a2.testpolymod1)

a2.test.pred.poly <- predict(a2.testpolymod1,traintest.data) 
 
a2.RMSE.polymod <- sqrt(mean((a2.test.pred.poly-traintest.data$a2)^2))
a2.RMSE.polymod

a2.MAE.polymod <- mean(abs(a2.test.pred.poly-traintest.data$a2))
a2.MAE.polymod

# checking for lower polys

a2.testpolymod2 <- glm(a2 ~ poly(as.numeric(speed),1) + poly(Chla,2), data=newtrain.data)
summary(a2.testpolymod2)

a2.test.pred.poly2 <- predict(a2.testpolymod2,traintest.data) 
 
a2.RMSE.polymod2 <- sqrt(mean((a2.test.pred.poly2-traintest.data$a2)^2))
a2.RMSE.polymod2

a2.MAE.polymod2 <- mean(abs(a2.test.pred.poly2-traintest.data$a2))
a2.MAE.polymod2

# Best poly is a2.testpolymod2, alternatively just Chla

a2.testpolymod3 <- glm(a2 ~ poly(Chla,2), data=newtrain.data)
summary(a2.testpolymod3)

a2.test.pred.poly3 <- predict(a2.testpolymod3,traintest.data) 
 
a2.RMSE.polymod3 <- sqrt(mean((a2.test.pred.poly3-traintest.data$a2)^2))
a2.RMSE.polymod3

a2.MAE.polymod3 <- mean(abs(a2.test.pred.poly3-traintest.data$a2))
a2.MAE.polymod3

# slightly worse

```


```{r}

## plot the rmse
a2.models.rmse <- tibble(
            a2.model = paste0("model.pd",c(1,2,3)),
						RMSE= c(
							c.rmse(traintest.data$a2,predict(a2.testpolymod1,traintest.data)),
              c.rmse(traintest.data$a2,predict(a2.testpolymod2,traintest.data)),
							c.rmse(traintest.data$a2,predict(a2.testpolymod3,traintest.data))
							)
					)
a2.models.rmse
a2.ncoef <- function(a2.model){
	a2.model %>%
    coefficients %>%
    length %>%
    {. - 1}
}


a2.models.rmse$a2.ncoef <- c(a2.ncoef(a2.testpolymod1),
                        a2.ncoef(a2.testpolymod2),
                        a2.ncoef(a2.testpolymod3))

a2.models.rmse %>%
  ggplot(aes(x=a2.ncoef, y= RMSE)) +
  geom_line(color = "dodgerblue") + 
  geom_point(color = "dodgerblue") +
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10) )

# plot the fit

a2.fitp1 <- lm(a2 ~ poly(as.numeric(speed),1) + poly(Chla,2), data=train.data )
summary(a2.fitp1)

a2.fitp2 <- lm(a2 ~ poly(as.numeric(speed),2) + poly(Chla,2), data=train.data )
summary(a2.fitp2)

a2.fitp3<- lm(a2 ~ poly(Chla,2), data=train.data)
summary(a2.fitp3)

train.data <- train.data %>%
	mutate(fit1 = predict(a2.fitp1),
	fit2 = predict(a2.fitp2),
	fit3 = predict(a2.fitp3))

#  visualization: doesnt make so much sense, but at least having it visualized / too many variables

cols <- c( "Deg.2", "Deg.1")
train.data %>% 
	ggplot(aes(x=a2, y=Chla)) +
	geom_point() +
	geom_line(aes(y=fit1, color="deg 3"), size =1) +
	geom_line(aes(y=fit2, color="deg 4"), size =1) +
	geom_line(aes(y=fit3, color="deg 2"), size =1) +
  theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```


## Log Linear Model

```{r}

a2.testlogmod <- lm(log(a2+1) ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a2.testlogmod)

exp(a2.testlogmod$coefficients)

a2.test.pred.log <- (predict(a2.testlogmod,traintest.data))
 
a2.RMSE.testlog <- sqrt(mean((a2.test.pred.log-traintest.data$a2)^2))
a2.RMSE.testlog

a2.MAE.testlog <- mean(abs(a2.test.pred.log-traintest.data$a2))
a2.MAE.testlog

# deleting the totally unsignificant

a2.testlogmod1 <- lm(log(a2+1) ~ mxPH + Chla , data=newtrain.data)
summary(a2.testlogmod1)
a2.testlogmod1$coefficients

exp(a2.testlogmod1$coefficients)

a2.test.pred.log2 <- (predict(a2.testlogmod1,traintest.data))
 
a2.RMSE.testlog2 <- sqrt(mean((a2.test.pred.log2-traintest.data$a2)^2))
a2.RMSE.testlog2

a2.MAE.testlog2 <- mean(abs(a2.test.pred.log2-traintest.data$a2))
a2.MAE.testlog2

# not increasing, a2.testlogmod better

```



## Trees

```{r}
require(rpart)
require(rattle)
# prepare the testing

a2.testtreemod <- rpart(a2 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data,method="anova")

a2.test.pred.treemod <- predict(a2.testtreemod,traintest.data) 
 
a2.RMSE.treemod <- sqrt(mean((a2.test.pred.treemod-traintest.data$a2)^2))
a2.RMSE.treemod

a2.MAE.treemod <- mean(abs(a2.test.pred.treemod-traintest.data$a2))
a2.MAE.treemod

# pruning it

printcp(a2.testtreemod)

min.xerror <- a2.testtreemod$cptable[which.min(a2.testtreemod$cptable[,"xerror"]),"CP"]

a2.treemod <- prune(a2.testtreemod, cp = min.xerror)
summary(a2.treemod)

a2.test.pred.treemod2 <- predict(a2.treemod,traintest.data) 
 
a2.RMSE.treemod2 <- sqrt(mean((a2.test.pred.treemod2-traintest.data$a2)^2))
a2.RMSE.treemod2

a2.MAE.treemod2 <- mean(abs(a2.test.pred.treemod2-traintest.data$a2))
a2.MAE.treemod2

plot(a2.treemod)
text(a2.treemod, pretty = 0)

```

## RandomForest

```{r}
# test it
require(randomForest)
require(rattle)
set.seed(200)
a2.forestmod <- randomForest(a2 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data, na.action = na.omit, importance = TRUE, ntree=1000)

which.min(a2.forestmod$mse)


print(a2.forestmod)
plot(a2.forestmod)

# understanding the importance of each variable
# a2.forestmodimp <- as.data.frame(sort(importance(a2.forestmod)[,1],decreasing = TRUE),optional = T)
# names(a2.forestmodimp) <- "% Inc MSE"
# a2.forestmodimp
a2.forestmod$importance

a2.testforest <- predict(a2.forestmod,traintest.data)
a2.RMSE.forestmod <- sqrt(mean((a2.testforest-traintest.data$a2)^2))
a2.RMSE.forestmod
 
a2.MAE.forestmod <- mean(abs(a2.testforest-traintest.data$a2))
a2.MAE.forestmod
```

```{r}
a2.accuracy <- data.frame(Method = c("MLR","Poly LR","Log LR","Pruned Tree","RandomForest"),
                         a2.RMSE   = c(a2.RMSE.testmlr4,a2.RMSE.polymod2,a2.RMSE.testlog,a2.RMSE.treemod2
,a2.RMSE.forestmod),
                         a2.MAE    = c(a2.MAE.testmlr4,a2.MAE.polymod2,a2.MAE.testlog,a2.MAE.treemod2,a2.MAE.forestmod)) 

a2.accuracy

```

The table shows that again RandomForest is the best method for a2.

Now we make the prediction:

```{r}
a2.to.pred <-randomForest(a2 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = train.data, na.action = na.omit, importance = TRUE, ntree=1000)

a2.pred <- predict(a2.to.pred, test.data)
a2.pred
# all negative values to zero
a2.pred[a2.pred<0] <- 0
summary(a2.pred)
# just to have a comparison
summary(train.data$a2)
```