---
title: "12c_Solution_a3"
author: "CTrierweiler,PGaulke"
date: "29 Juli 2019"
output: html_document
---

# Develop the Model for a3 

Methods that will be tested:

- Multiple Linear Regression
- Linear regression with polynomials
- Log Linear Regression
- Trees
- RandomForest

Logistic regression is not applicable on this data set as the response is not binary.

## Multiple Linear Regression

```{r}

a3.testmlrmod <- lm(a3 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a3.testmlrmod)

a3.test.pred.mlr <- (predict(a3.testmlrmod,traintest.data))
 
a3.RMSE.testmlr <- sqrt(mean((a3.test.pred.mlr-traintest.data$a3)^2))
a3.RMSE.testmlr

a3.MAE.testmlr <- mean(abs(a3.test.pred.mlr-traintest.data$a3))
a3.MAE.testmlr

# Reducing the variables
a3.testmlrmod2 <- lm(a3 ~ NH4 + mnO2 + season, data=newtrain.data)
summary(a3.testmlrmod2)

a3.test.pred.mlr2 <- (predict(a3.testmlrmod2,traintest.data))
 
a3.RMSE.testmlr2 <- sqrt(mean((a3.test.pred.mlr2-traintest.data$a3)^2))
a3.RMSE.testmlr2

a3.MAE.testmlr2 <- mean(abs(a3.test.pred.mlr2-traintest.data$a3))
a3.MAE.testmlr2

# Reducing the variables to only one variable (SLR)

a3.testmlrmod3 <- lm(a3 ~ season, data=newtrain.data)

summary(a3.testmlrmod3)

a3.test.pred.mlr3 <- (predict(a3.testmlrmod3,traintest.data))
 
a3.RMSE.testmlr3 <- sqrt(mean((a3.test.pred.mlr3-traintest.data$a3)^2))
a3.RMSE.testmlr3

a3.MAE.testmlr3 <- mean(abs(a3.test.pred.mlr3-traintest.data$a3))
a3.MAE.testmlr3

# SLR best, optimize
a3.testmlrmod4 <- lm(a3 ~ season + NH4 * mnO2, data=newtrain.data)
summary(a3.testmlrmod4)
a3.test.pred.mlr4 <- (predict(a3.testmlrmod4,traintest.data))
 
a3.RMSE.testmlr4 <- sqrt(mean((a3.test.pred.mlr4-traintest.data$a3)^2))
a3.RMSE.testmlr4

a3.MAE.testmlr4 <- mean(abs(a3.test.pred.mlr4-traintest.data$a3))
a3.MAE.testmlr4

# a3-testmlrmod4 as best choice

```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a3.crossmodelmlr <- train(a3 ~ season + NH4 * mnO2, data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a3.crossmodelmlr)
```


## Polynomials

Adding polynomials

```{r}

# test it from mlr4

a3.testpoly <- glm(a3 ~ season + NH4 * mnO2, data=newtrain.data)
summary(a3.testpoly)
# use of poly

a3.testpolymod1 <- glm(a3 ~ poly(as.numeric(season),2) + poly(NH4,2) * poly(mnO2,2), data=newtrain.data)
summary(a3.testpolymod1)

a3.test.pred.poly <- predict(a3.testpolymod1,traintest.data) 
 
a3.RMSE.polymod <- sqrt(mean((a3.test.pred.poly-traintest.data$a3)^2))
a3.RMSE.polymod

a3.MAE.polymod <- mean(abs(a3.test.pred.poly-traintest.data$a3))
a3.MAE.polymod

# checking for adjusted polys

a3.testpolymod2 <- glm(a3 ~ poly(as.numeric(season),3) + NH4 * mnO2, data=newtrain.data)
summary(a3.testpolymod2)

a3.test.pred.poly2 <- predict(a3.testpolymod2,traintest.data) 
 
a3.RMSE.polymod2 <- sqrt(mean((a3.test.pred.poly2-traintest.data$a3)^2))
a3.RMSE.polymod2

a3.MAE.polymod2 <- mean(abs(a3.test.pred.poly2-traintest.data$a3))
a3.MAE.polymod2

# Best poly is a3.testpolymod2

a3.testpolymod3 <- glm(a3 ~ poly(as.numeric(season),3), data=newtrain.data)
summary(a3.testpolymod3)

a3.test.pred.poly3 <- predict(a3.testpolymod3,traintest.data) 
 
a3.RMSE.polymod3 <- sqrt(mean((a3.test.pred.poly3-traintest.data$a3)^2))
a3.RMSE.polymod3

a3.MAE.polymod3 <- mean(abs(a3.test.pred.poly3-traintest.data$a3))
a3.MAE.polymod3

# worse
```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a3.crossmodelmlr <- train(a3 ~ poly(as.numeric(season),3) + NH4 * mnO2, data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a3.crossmodelmlr)
```

```{r}
## plot the rmse
a3.models.rmse <- tibble(
            a3.model = paste0("model.pd",c(1,2,3)),
						RMSE= c(
							c.rmse(traintest.data$a3,predict(a3.testpolymod1,traintest.data)),
              c.rmse(traintest.data$a3,predict(a3.testpolymod2,traintest.data)),
							c.rmse(traintest.data$a3,predict(a3.testpolymod3,traintest.data))
							)
					)
a3.models.rmse
a3.ncoef <- function(a3.model){
	a3.model %>%
    coefficients %>%
    length %>%
    {. - 1}
}


a3.models.rmse$a3.ncoef <- c(a3.ncoef(a3.testpolymod1),
                        a3.ncoef(a3.testpolymod2),
                        a3.ncoef(a3.testpolymod3))

a3.models.rmse %>%
  ggplot(aes(x=a3.ncoef, y= RMSE)) +
  geom_line(color = "dodgerblue") + 
  geom_point(color = "dodgerblue") +
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10) )

# plot the fit

a3.fitp1 <- lm(a3 ~ poly(as.numeric(season),3) + NH4 * mnO2, data=train.data )
summary(a3.fitp1)

a3.fitp2 <- lm(a3 ~ poly(as.numeric(season),2) + poly(NH4,2) * poly(mnO2,2), data=train.data )
summary(a3.fitp2)

a3.fitp3 <- lm(a3 ~ poly(as.numeric(season),3), data=train.data )
summary(a3.fitp3)

train.data <- train.data %>%
	mutate(fit1 = predict(a3.fitp1),
	fit2 = predict(a3.fitp2),
	fit3 = predict(a3.fitp3))

#  visualization: doesnt make so much sense, but at least having it visualized / too many variables

cols <- c( "Deg.2", "Deg.1")
train.data %>% 
	ggplot(aes(x=a3, y=PO4)) +
	geom_point() +
	geom_line(aes(y=fit1, color="deg 5"), size =1) +
	geom_line(aes(y=fit2, color="deg 6"), size =1) +
  geom_line(aes(y=fit2, color="deg 3"), size =1) +

	theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```


## Log Linear Model

```{r}

a3.testlogmod <- lm(log(a3+1) ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a3.testlogmod)

exp(a3.testlogmod$coefficients)

a3.test.pred.log <- (predict(a3.testlogmod,traintest.data))
 
a3.RMSE.testlog <- sqrt(mean((a3.test.pred.log-traintest.data$a3)^2))
a3.RMSE.testlog

a3.MAE.testlog <- mean(abs(a3.test.pred.log-traintest.data$a3))
a3.MAE.testlog

# deleting the totally unsignificant

a3.testlogmod1 <- lm(log(a3+1) ~ season + NH4 * mnO2 , data=newtrain.data)
summary(a3.testlogmod1)
a3.testlogmod1$coefficients

exp(a3.testlogmod1$coefficients)

a3.test.pred.log2 <- (predict(a3.testlogmod1,traintest.data))
 
a3.RMSE.testlog2 <- sqrt(mean((a3.test.pred.log2-traintest.data$a3)^2))
a3.RMSE.testlog2

a3.MAE.testlog2 <- mean(abs(a3.test.pred.log2-traintest.data$a3))
a3.MAE.testlog2

# not increasing, a3.testlogmod better

```

## Trees

```{r}
require(rpart)
require(rattle)
# prepare the testing

a3.testtreemod <- rpart(a3 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data,method="anova")

a3.test.pred.treemod <- predict(a3.testtreemod,traintest.data) 
 
a3.RMSE.treemod <- sqrt(mean((a3.test.pred.treemod-traintest.data$a3)^2))
a3.RMSE.treemod

a3.MAE.treemod <- mean(abs(a3.test.pred.treemod-traintest.data$a3))
a3.MAE.treemod

# pruning it

printcp(a3.testtreemod)

min.xerror <- a3.testtreemod$cptable[which.min(a3.testtreemod$cptable[,"xerror"]),"CP"]

a3.treemod <- prune(a3.testtreemod, cp = min.xerror)
summary(a3.treemod)

a3.test.pred.treemod2 <- predict(a3.treemod,traintest.data) 
 
a3.RMSE.treemod2 <- sqrt(mean((a3.test.pred.treemod2-traintest.data$a3)^2))
a3.RMSE.treemod2

a3.MAE.treemod2 <- mean(abs(a3.test.pred.treemod2-traintest.data$a3))
a3.MAE.treemod2

plot(a3.treemod)
text(a3.treemod, pretty = 0)

```

## RandomForest

```{r}
# test it
require(randomForest)
require(rattle)
set.seed(200)
a3.forestmod <- randomForest(a3 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data, na.action = na.omit, importance = TRUE, ntree=1000)

which.min(a3.forestmod$mse)


print(a3.forestmod)
plot(a3.forestmod)

# understanding the importance of each variable
# a3.forestmodimp <- as.data.frame(sort(importance(a3.forestmod)[,1],decreasing = TRUE),optional = T)
# names(a3.forestmodimp) <- "% Inc MSE"
# a3.forestmodimp
a3.forestmod$importance

a3.testforest <- predict(a3.forestmod,traintest.data)
a3.RMSE.forestmod <- sqrt(mean((a3.testforest-traintest.data$a3)^2))
a3.RMSE.forestmod
 
a3.MAE.forestmod <- mean(abs(a3.testforest-traintest.data$a3))
a3.MAE.forestmod
```

```{r}
a3.accuracy <- data.frame(Method = c("MLR","Poly LR","Log LR","Pruned Tree","RandomForest"),
                         a3.RMSE   = c(a3.RMSE.testmlr4,a3.RMSE.polymod2,a3.RMSE.testlog,a3.RMSE.treemod2
,a3.RMSE.forestmod),
                         a3.MAE    = c(a3.MAE.testmlr4,a3.MAE.polymod2,a3.MAE.testlog,a3.MAE.treemod2,a3.MAE.forestmod)) 

a3.accuracy

```

The table shows that MLR and Poly are the best method for a3.

Now we make the prediction(with a poly):

```{r}
a3.to.pred <-lm(a3 ~ poly(as.numeric(season),3) + NH4 * mnO2, data = train.data)

a3.pred <- predict(a3.to.pred, test.data)
a3.pred
# all negative values to zero
a3.pred[a3.pred<0] <- 0
summary(a3.pred)
# just to have a comparison
summary(train.data$a3)
```