---
title: "12e_Solution_a5"
author: "CTrierweiler,PGaulke"
date: "29 Juli 2019"
output: html_document
---

# Develop the Model for a5 

Methods that will be tested:

- Multiple Linear Regression
- Linear regression with polynomials
- Log Linear Regression
- Trees
- RandomForest

Logistic regression is not applicable on this data set as the response is not binary.

## Multiple Linear Regression

```{r}

a5.testmlrmod <- lm(a5 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a5.testmlrmod)

a5.test.pred.mlr <- (predict(a5.testmlrmod,traintest.data))
 
a5.RMSE.testmlr <- sqrt(mean((a5.test.pred.mlr-traintest.data$a5)^2))
a5.RMSE.testmlr

a5.MAE.testmlr <- mean(abs(a5.test.pred.mlr-traintest.data$a5))
a5.MAE.testmlr

# Reducing the variables
a5.testmlrmod2 <- lm(a5 ~ size *season + mnO2, data=newtrain.data)
summary(a5.testmlrmod2)

a5.test.pred.mlr2 <- (predict(a5.testmlrmod2,traintest.data))
 
a5.RMSE.testmlr2 <- sqrt(mean((a5.test.pred.mlr2-traintest.data$a5)^2))
a5.RMSE.testmlr2

a5.MAE.testmlr2 <- mean(abs(a5.test.pred.mlr2-traintest.data$a5))
a5.MAE.testmlr2

# Reducing the variables to only one variable (SLR)

a5.testmlrmod3 <- lm(a5 ~ mnO2, data=newtrain.data)

summary(a5.testmlrmod3)

a5.test.pred.mlr3 <- (predict(a5.testmlrmod3,traintest.data))
 
a5.RMSE.testmlr3 <- sqrt(mean((a5.test.pred.mlr3-traintest.data$a5)^2))
a5.RMSE.testmlr3

a5.MAE.testmlr3 <- mean(abs(a5.test.pred.mlr3-traintest.data$a5))
a5.MAE.testmlr3

# SLR best, optimize
a5.testmlrmod4 <- lm(a5 ~ size * mnO2, data=newtrain.data)
summary(a5.testmlrmod4)
a5.test.pred.mlr4 <- (predict(a5.testmlrmod4,traintest.data))
 
a5.RMSE.testmlr4 <- sqrt(mean((a5.test.pred.mlr4-traintest.data$a5)^2))
a5.RMSE.testmlr4

a5.MAE.testmlr4 <- mean(abs(a5.test.pred.mlr4-traintest.data$a5))
a5.MAE.testmlr4

# a5-testmlrmod4 as best choice

```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a5.crossmodelmlr <- train(a5 ~ size * mnO2, data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a5.crossmodelmlr)
```


## Polynomials

Adding polynomials

```{r}

# test it from mlr4

a5.testpoly <- glm(a5 ~ size * mnO2, data=newtrain.data)
summary(a5.testpoly)
# use of poly

a5.testpolymod1 <- glm(a5 ~ poly(as.numeric(size),2) * poly(mnO2,2), data=newtrain.data)
summary(a5.testpolymod1)

a5.test.pred.poly <- predict(a5.testpolymod1,traintest.data) 
 
a5.RMSE.polymod <- sqrt(mean((a5.test.pred.poly-traintest.data$a5)^2))
a5.RMSE.polymod

a5.MAE.polymod <- mean(abs(a5.test.pred.poly-traintest.data$a5))
a5.MAE.polymod

# checking for adjusted polys

a5.testpolymod2 <- glm(a5 ~ poly(as.numeric(size),2) * poly(mnO2,1), data=newtrain.data)
summary(a5.testpolymod2)

a5.test.pred.poly2 <- predict(a5.testpolymod2,traintest.data) 
 
a5.RMSE.polymod2 <- sqrt(mean((a5.test.pred.poly2-traintest.data$a5)^2))
a5.RMSE.polymod2

a5.MAE.polymod2 <- mean(abs(a5.test.pred.poly2-traintest.data$a5))
a5.MAE.polymod2

# Best poly is a5.testpolymod2, alternative

a5.testpolymod3 <- glm(a5 ~ poly(as.numeric(size),2), data=newtrain.data)
summary(a5.testpolymod3)

a5.test.pred.poly3 <- predict(a5.testpolymod3,traintest.data) 
 
a5.RMSE.polymod3 <- sqrt(mean((a5.test.pred.poly3-traintest.data$a5)^2))
a5.RMSE.polymod3

a5.MAE.polymod3 <- mean(abs(a5.test.pred.poly3-traintest.data$a5))
a5.MAE.polymod3

# worse

```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a5.crossmodelmlr <- train(a5 ~ poly(as.numeric(size),2) * poly(mnO2,1), data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a5.crossmodelmlr)
```

```{r}
## plot the rmse
a5.models.rmse <- tibble(
            a5.model = paste0("model.pd",c(1,2,3)),
						RMSE= c(
							c.rmse(traintest.data$a5,predict(a5.testpolymod1,traintest.data)),
              c.rmse(traintest.data$a5,predict(a5.testpolymod2,traintest.data)),
							c.rmse(traintest.data$a5,predict(a5.testpolymod3,traintest.data))
							)
					)
a5.models.rmse
a5.ncoef <- function(a5.model){
	a5.model %>%
    coefficients %>%
    length %>%
    {. - 1}
}


a5.models.rmse$a5.ncoef <- c(a5.ncoef(a5.testpolymod1),
                        a5.ncoef(a5.testpolymod2),
                        a5.ncoef(a5.testpolymod3))

a5.models.rmse %>%
  ggplot(aes(x=a5.ncoef, y= RMSE)) +
  geom_line(color = "dodgerblue") + 
  geom_point(color = "dodgerblue") +
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10) )

# plot the fit

a5.fitp1 <- lm(a5 ~ poly(as.numeric(size),2) * poly(mnO2,2), data=train.data )
summary(a5.fitp1)

a5.fitp2 <- lm(a5 ~ poly(as.numeric(size),2) * poly(mnO2,1), data=train.data )
summary(a5.fitp2)

a5.fitp3 <- lm(a5 ~ poly(as.numeric(size),2), data=train.data )
summary(a5.fitp3)


train.data <- train.data %>%
	mutate(fit1 = predict(a5.fitp1),
	fit2 = predict(a5.fitp2),
	fit3 = predict(a5.fitp3))

#  visualization: doesnt make so much sense, but at least having it visualized / too many variables

cols <- c( "Deg.2", "Deg.1")
train.data %>% 
	ggplot(aes(x=a5, y=size)) +
	geom_point() +
	geom_line(aes(y=fit1, color="deg 8"), size =1) +
	geom_line(aes(y=fit2, color="deg 3"), size =1) +
  geom_line(aes(y=fit3, color="deg 2"), size =1) +
  	theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```


## Log Linear Model

```{r}

a5.testlogmod <- lm(log(a5+1) ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a5.testlogmod)

exp(a5.testlogmod$coefficients)


a5.test.pred.log <- (predict(a5.testlogmod,traintest.data))
 
a5.RMSE.testlog <- sqrt(mean((a5.test.pred.log-traintest.data$a5)^2))
a5.RMSE.testlog

a5.MAE.testlog <- mean(abs(a5.test.pred.log-traintest.data$a5))
a5.MAE.testlog

# deleting the totally unsignificant

a5.testlogmod1 <- lm(log(a5+1) ~ size *mnO2 , data=newtrain.data)
summary(a5.testlogmod1)
a5.testlogmod1$coefficients

exp(a5.testlogmod1$coefficients)

a5.test.pred.log2 <- (predict(a5.testlogmod1,traintest.data))
 
a5.RMSE.testlog2 <- sqrt(mean((a5.test.pred.log2-traintest.data$a5)^2))
a5.RMSE.testlog2

a5.MAE.testlog2 <- mean(abs(a5.test.pred.log2-traintest.data$a5))
a5.MAE.testlog2

# not increasing, a5.testlogmod better

```

## Trees

```{r, error=FALSE}
require(rpart)
require(rattle)
# prepare the testing

a5.testtreemod <- rpart(a5 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data,method="anova")

a5.test.pred.treemod <- predict(a5.testtreemod,traintest.data) 
 
a5.RMSE.treemod <- sqrt(mean((a5.test.pred.treemod-traintest.data$a5)^2))
a5.RMSE.treemod

a5.MAE.treemod <- mean(abs(a5.test.pred.treemod-traintest.data$a5))
a5.MAE.treemod

# pruning it

printcp(a5.testtreemod)

min.xerror <- a5.testtreemod$cptable[which.min(a5.testtreemod$cptable[,"xerror"]),"CP"]

a5.treemod <- prune(a5.testtreemod, cp = min.xerror)
summary(a5.treemod)

a5.test.pred.treemod2 <- predict(a5.treemod,traintest.data) 
 
a5.RMSE.treemod2 <- sqrt(mean((a5.test.pred.treemod2-traintest.data$a5)^2))
a5.RMSE.treemod2

a5.MAE.treemod2 <- mean(abs(a5.test.pred.treemod2-traintest.data$a5))
a5.MAE.treemod2

# plot(a5.treemod)
# text(a5.treemod, pretty = 0)

```

## RandomForest

```{r}
# test it
require(randomForest)
require(rattle)
set.seed(200)
a5.forestmod <- randomForest(a5 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data, na.action = na.omit, importance = TRUE, ntree=1000)

which.min(a5.forestmod$mse)


print(a5.forestmod)
plot(a5.forestmod)

# understanding the importance of each variable
# a5.forestmodimp <- as.data.frame(sort(importance(a5.forestmod)[,1],decreasing = TRUE),optional = T)
# names(a5.forestmodimp) <- "% Inc MSE"
# a5.forestmodimp
a5.forestmod$importance

a5.testforest <- predict(a5.forestmod,traintest.data)
a5.RMSE.forestmod <- sqrt(mean((a5.testforest-traintest.data$a5)^2))
a5.RMSE.forestmod
 
a5.MAE.forestmod <- mean(abs(a5.testforest-traintest.data$a5))
a5.MAE.forestmod
```

```{r}
a5.accuracy <- data.frame(Method = c("MLR","Poly LR","Log LR","Pruned Tree","RandomForest"),
                         a5.RMSE   = c(a5.RMSE.testmlr4,a5.RMSE.polymod2,a5.RMSE.testlog,a5.RMSE.treemod2
,a5.RMSE.forestmod),
                         a5.MAE    = c(a5.MAE.testmlr4,a5.MAE.polymod2,a5.MAE.testlog,a5.MAE.treemod2,a5.MAE.forestmod)) 

a5.accuracy

```

The table shows that Random Forest is best method for a5.

Now we make the prediction:

```{r}
a5.to.pred <-randomForest(a5 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data, na.action = na.omit, importance = TRUE, ntree=1000)

a5.pred <- predict(a5.to.pred, test.data)
a5.pred
# all negative values to zero
a5.pred[a5.pred<0] <- 0
summary(a5.pred)
# just to have a comparison
summary(train.data$a5)
```