---
title: "12g_Solution_a7"
author: "CTrierweiler,PGaulke"
date: "29 Juli 2019"
output: html_document
---
# Develop the Model for a7 

Methods that will be tested:

- Multiple Linear Regression
- Linear regression with polynomials
- Log Linear Regression
- Trees
- RandomForest

Logistic regression is not applicable on this data set as the response is not binary.

## Multiple Linear Regression

```{r}

a7.testmlrmod <- lm(a7 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a7.testmlrmod)

a7.test.pred.mlr <- (predict(a7.testmlrmod,traintest.data))
 
a7.RMSE.testmlr <- sqrt(mean((a7.test.pred.mlr-traintest.data$a7)^2))
a7.RMSE.testmlr

a7.MAE.testmlr <- mean(abs(a7.test.pred.mlr-traintest.data$a7))
a7.MAE.testmlr

# Reducing the variables
a7.testmlrmod2 <- lm(a7 ~ Chla + NO3, data=newtrain.data)
summary(a7.testmlrmod2)

a7.test.pred.mlr2 <- (predict(a7.testmlrmod2,traintest.data))
 
a7.RMSE.testmlr2 <- sqrt(mean((a7.test.pred.mlr2-traintest.data$a7)^2))
a7.RMSE.testmlr2

a7.MAE.testmlr2 <- mean(abs(a7.test.pred.mlr2-traintest.data$a7))
a7.MAE.testmlr2

# Reducing the variables to only one variable (SLR)

a7.testmlrmod3 <- lm(a7 ~ NO3, data=newtrain.data)

summary(a7.testmlrmod3)

a7.test.pred.mlr3 <- (predict(a7.testmlrmod3,traintest.data))
 
a7.RMSE.testmlr3 <- sqrt(mean((a7.test.pred.mlr3-traintest.data$a7)^2))
a7.RMSE.testmlr3

a7.MAE.testmlr3 <- mean(abs(a7.test.pred.mlr3-traintest.data$a7))
a7.MAE.testmlr3

# SLR best, optimize
a7.testmlrmod4 <- lm(a7 ~  NO3 * Chla, data=newtrain.data)
summary(a7.testmlrmod4)
a7.test.pred.mlr4 <- (predict(a7.testmlrmod4,traintest.data))
 
a7.RMSE.testmlr4 <- sqrt(mean((a7.test.pred.mlr4-traintest.data$a7)^2))
a7.RMSE.testmlr4

a7.MAE.testmlr4 <- mean(abs(a7.test.pred.mlr4-traintest.data$a7))
a7.MAE.testmlr4

# a7-testmlrmod4 as best choice

```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a7.crossmodelmlr <- train(a7 ~ NO3 * Chla, data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a7.crossmodelmlr)
```


## Polynomials

Adding polynomials

```{r}

# test it from mlr4

a7.testpoly <- glm(a7 ~ NO3 * Chla, data=newtrain.data)
summary(a7.testpoly)
# use of poly

a7.testpolymod1 <- glm(a7 ~ poly(NO3,2) * poly(Chla,2), data=newtrain.data)
summary(a7.testpolymod1)

a7.test.pred.poly <- predict(a7.testpolymod1,traintest.data) 
 
a7.RMSE.polymod <- sqrt(mean((a7.test.pred.poly-traintest.data$a7)^2))
a7.RMSE.polymod

a7.MAE.polymod <- mean(abs(a7.test.pred.poly-traintest.data$a7))
a7.MAE.polymod

# checking for adjusted polys

a7.testpolymod2 <- glm(a7 ~ poly(NO3,1) * poly(Chla,2), data=newtrain.data)
summary(a7.testpolymod2)

a7.test.pred.poly2 <- predict(a7.testpolymod2,traintest.data) 
 
a7.RMSE.polymod2 <- sqrt(mean((a7.test.pred.poly2-traintest.data$a7)^2))
a7.RMSE.polymod2

a7.MAE.polymod2 <- mean(abs(a7.test.pred.poly2-traintest.data$a7))
a7.MAE.polymod2

# Best poly is a7.testpolymod2

a7.testpolymod3 <- glm(a7 ~ poly(NO3,2), data=newtrain.data)
summary(a7.testpolymod3)

a7.test.pred.poly3 <- predict(a7.testpolymod3,traintest.data) 
 
a7.RMSE.polymod3 <- sqrt(mean((a7.test.pred.poly3-traintest.data$a7)^2))
a7.RMSE.polymod3

a7.MAE.polymod3 <- mean(abs(a7.test.pred.poly3-traintest.data$a7))
a7.MAE.polymod3


# worse

```

Trying k-fold cross validation

```{r}
require(caret)
set.seed(200) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
a7.crossmodelmlr <- train(a7 ~ poly(NO3,1) * poly(Chla,2), data = train.data, method = "lm",
               trControl = train.control)
# Summarize the results
print(a7.crossmodelmlr)
```

```{r}
## plot the rmse
a7.models.rmse <- tibble(
            a7.model = paste0("model.pd",c(1,2,3)),
						RMSE= c(
							c.rmse(traintest.data$a7,predict(a7.testpolymod1,traintest.data)),
              c.rmse(traintest.data$a7,predict(a7.testpolymod2,traintest.data)),
							c.rmse(traintest.data$a7,predict(a7.testpolymod3,traintest.data))
							)
					)
a7.models.rmse
a7.ncoef <- function(a7.model){
	a7.model %>%
    coefficients %>%
    length %>%
    {. - 1}
}


a7.models.rmse$a7.ncoef <- c(a7.ncoef(a7.testpolymod1),
                        a7.ncoef(a7.testpolymod2),
                        a7.ncoef(a7.testpolymod3))

a7.models.rmse %>%
  ggplot(aes(x=a7.ncoef, y= RMSE)) +
  geom_line(color = "dodgerblue") + 
  geom_point(color = "dodgerblue") +
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10) )

# plot the fit

a7.fitp1 <- lm(a7 ~ poly(NO3,1) * poly(Chla,2), data=train.data )
summary(a7.fitp1)

a7.fitp2 <- lm(a7 ~ poly(NO3,2) * poly(Chla,2), data=train.data )
summary(a7.fitp2)


a7.fitp3 <- lm(a7 ~ poly(NO3,2), data=train.data )
summary(a7.fitp3)

train.data <- train.data %>%
	mutate(fit1 = predict(a7.fitp1),
	fit2 = predict(a7.fitp2),
	fit3 = predict(a7.fitp3))

#  visualization: doesnt make so much sense, but at least having it visualized / too many variables

cols <- c( "Deg.2", "Deg.1")
train.data %>% 
	ggplot(aes(x=a7, y=PO4)) +
	geom_point() +
	geom_line(aes(y=fit1, color="deg 8"), size =1) +
	geom_line(aes(y=fit2, color="deg 5"), size =1) +
	geom_line(aes(y=fit3, color="deg 2"), size =1) +
  theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```


## Log Linear Model

```{r}

a7.testlogmod <- lm(log(a7+1) ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data=newtrain.data)
summary(a7.testlogmod)

exp(a7.testlogmod$coefficients)

a7.test.pred.log <- (predict(a7.testlogmod,traintest.data))
 
a7.RMSE.testlog <- sqrt(mean((a7.test.pred.log-traintest.data$a7)^2))
a7.RMSE.testlog

a7.MAE.testlog <- mean(abs(a7.test.pred.log-traintest.data$a7))
a7.MAE.testlog

# deleting the totally unsignificant

a7.testlogmod1 <- lm(log(a7+1) ~ season*Chla*Cl , data=newtrain.data)
summary(a7.testlogmod1)
a7.testlogmod1$coefficients

exp(a7.testlogmod1$coefficients)

a7.test.pred.log2 <- (predict(a7.testlogmod1,traintest.data))
 
a7.RMSE.testlog2 <- sqrt(mean((a7.test.pred.log2-traintest.data$a7)^2))
a7.RMSE.testlog2

a7.MAE.testlog2 <- mean(abs(a7.test.pred.log2-traintest.data$a7))
a7.MAE.testlog2

# not increasing, a7.testlogmod better

```

## Trees

```{r, error=FALSE}
require(rpart)
require(rattle)
# prepare the testing

a7.testtreemod <- rpart(a7 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data,method="anova")

a7.test.pred.treemod <- predict(a7.testtreemod,traintest.data) 
 
a7.RMSE.treemod <- sqrt(mean((a7.test.pred.treemod-traintest.data$a7)^2))
a7.RMSE.treemod

a7.MAE.treemod <- mean(abs(a7.test.pred.treemod-traintest.data$a7))
a7.MAE.treemod

# pruning it

printcp(a7.testtreemod)

min.xerror <- a7.testtreemod$cptable[which.min(a7.testtreemod$cptable[,"xerror"]),"CP"]

a7.treemod <- prune(a7.testtreemod, cp = min.xerror)
summary(a7.treemod)

a7.test.pred.treemod2 <- predict(a7.treemod,traintest.data) 
 
a7.RMSE.treemod2 <- sqrt(mean((a7.test.pred.treemod2-traintest.data$a7)^2))
a7.RMSE.treemod2

a7.MAE.treemod2 <- mean(abs(a7.test.pred.treemod2-traintest.data$a7))
a7.MAE.treemod2

# plot(a7.treemod)
# text(a7.treemod, pretty = 0)

```

## RandomForest

```{r}
# test it
require(randomForest)
require(rattle)
set.seed(200)
a7.forestmod <- randomForest(a7 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = newtrain.data, na.action = na.omit, importance = TRUE, ntree=1000)

which.min(a7.forestmod$mse)


print(a7.forestmod)
plot(a7.forestmod)

# understanding the importance of each variable
# a7.forestmodimp <- as.data.frame(sort(importance(a7.forestmod)[,1],decreasing = TRUE),optional = T)
# names(a7.forestmodimp) <- "% Inc MSE"
# a7.forestmodimp
a7.forestmod$importance

a7.testforest <- predict(a7.forestmod,traintest.data)
a7.RMSE.forestmod <- sqrt(mean((a7.testforest-traintest.data$a7)^2))
a7.RMSE.forestmod
 
a7.MAE.forestmod <- mean(abs(a7.testforest-traintest.data$a7))
a7.MAE.forestmod
```

```{r}
a7.accuracy <- data.frame(Method = c("MLR","Poly LR","Log LR","Pruned Tree","RandomForest"),
                         a7.RMSE   = c(a7.RMSE.testmlr4,a7.RMSE.polymod2,a7.RMSE.testlog,a7.RMSE.treemod2
,a7.RMSE.forestmod),
                         a7.MAE    = c(a7.MAE.testmlr4,a7.MAE.polymod2,a7.MAE.testlog,a7.MAE.treemod2,a7.MAE.forestmod)) 

a7.accuracy

```

The table shows that RandomForest is the best method for a7.

Now we make the prediction:

```{r}
a7.to.pred <-randomForest(a7 ~ season + size + speed + mxPH + mnO2 + Cl + NO3 + NH4 + oPO4 + PO4 + Chla, data = train.data, na.action = na.omit, importance = TRUE, ntree=1000)

a7.pred <- predict(a7.to.pred, test.data)
a7.pred
# all negative values to zero
a7.pred[a7.pred<0] <- 0
summary(a7.pred)
# just to have a comparison
summary(train.data$a7)
```