---
title: "First Markdown"
author: "Philipp Gaulke"
date: 'Last update: `r format(Sys.time(), "%H:%m, %B %d, %Y")`'
output:
  pdf_document:
    highlight: "tango"
    number_sections: yes
    toc: no
  html_document:
    toc: no
  html_notebook:
    code_folding: hide
    highlight: tango
    number_sections: no
    theme: cerulean
    toc: no
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  word_document:
    toc: no
    toc_depth: '2'
---

#First Lesson


##How to Add Something in a R Markdown File

Two Ways
* written text, inclduing code which start with one backstick and ends with one backstick
* a Chunk: seperate field (see the grew field), it starts with 3 backsticks + r in brackets and ends with 3 backsticks


Further Options:
*use echo=false if you do not want to show it in the output file, if you say echo=true the chunk will be displayed
*if you ad eval=False, then the chunk will be ignored and the code in the written text will ignore the chunk

<<<<<<< HEAD:07-First_Steps.Rmd
* Header: YAML


# Forth Lesson 

```{r}
roots <- function(a, b, c){
  
  if (b^2- 4*a*c <0) {
    print("No solution! (negativ root cannot be squared)")
  } else {}
  
  (-b + sqrt (b^2- 4*a*c)) / 2*a
}
roots(a=5, b=6, c=1)
```
``` {r}
dna <- c ("A", "C", "G", "T")
a <- rep (dna, 12)
a
```




```{r}
library(tidyverse)
t_iris <- tibble(iris)
class(t_iris)
```

```{r}
my_data <- c(2:10, 20, 42)

library(magrittr)
my_data %>%
  mean () %>%
  sqrt () %>%
  round (2)


```


```{r}
my_data <- c(6:12, 23, 41)

library(magrittr)
my_data %>%
  mean () %>%
  sqrt () %>%
  round (2)


```


```{r}
library(tidyverse)


data("iris")
t_iris <- as_tibble(iris)
t_iris2 <- t_iris [t_iris$sepal.lenght>4.5 & t_iris$sepal.length<5.5,]

iris %>%
  as_tibble
```
=======
##How to Change the Header
Headerlanguage: YAML
>>>>>>> 2380010ebad2b5d0f1d845bd9cd8145df6bc82b6:01-First_Lesson.Rmd
