---
title: "10_Fourth_Lesson"
author: "PGaulke"
date: "5 April 2019"
output: pdf_document
---

#Fourth Lesson


####2.7 For the following questions, you may want to check the help on the function seq.
Create the following vectors:
• a = (1, 2, 3, ..., 15),
• b = (15, 14, 13, ..., 0,−1,−2),
• c = (1, 1.2, 1.4, ..., 8).
Create the following vectors, potentially based on auxiliary vectors.
• d = (0.531.31, 0.561.34, ..., 0.5361.334),
• e = (5, 52, 2 , 53, 3 , ..., 525
25 ).
Use the function all.equal to check that the last elements of each vector are indeed 1.0889154 × 10−7 and
1.1920929 × 1016, respectively.


```{r}
a <- seq(1,15,1)
a

b <- c(15:-2)

x0=1
c <- seq(x0,8,0.2)
c

d <- c(0.5^(seq(3,36,3))*1.3^seq(1,34,3))
d

xtfrm


```

####2.8Use paste in vector creation
For the following questions, you may want to check the help on the function paste.
Create the following vectors:
• a = (City 1, City 2, . . . , City 6),
• b = (n1, n2, . . . , n6)
• b = (Gender=M, Gender=F, Gender=M, Gender=F, Gender=M, Gender=F)
3


```{r}




```


##What happens in the lesson:

Differences between matrice and normal vectors. How we "access" or "subset" a matrice or vector: Vector just a list, Matrice has different dimension. Therefore, we need: mat.(row, column) to display what we want.
Shorter writing style for getting all values of a column: matrice$column
Than you can subset this matrice.


##How to get a subset out of a matrice
(BTW r condition has no logic, it is more a name)
```{r condition}
mtcars
mtcars[mtcars$mpg<20,] # all cars with mpg smaller 20


```

Further on vectors, put in conditions to display something like a statement

```{r}

gains <- c(10,3,-5,-4,12,4)
sum(gains)

# change the element -4 to -40 with a formula
# subset gains and assign new values
gains[gains==-4]<- -40
gains

if (sum(gains) >0) {
  print("Congratulations, you are winning")
} else  {
  print ("You are not winning")
}
# else statement for display something else in case the if condition is false

```


Less trivial ways of writing condition
  
```{r}
gains <- c(10,3,-5,0,-4,12,4)

if (is.numeric(gains)) {
  print ("Seems like it is a numeric vector...")
}

# here is-numeric (gains) evaluates to true hece the code is exceuted.

!is.data.frame(gains)

class(gains)

```


##Structures of simple function
###How to write functions

pow_two is just a random name here
```{r}
pow_two <- function (a){
  a^2
}
pow_two(12)
```

you can also here apply element by element evaluation


###Return of a function

```{r}
my_f2 <- function(a){
  a^2
  a^3 # this is the last exression executed
}
my_f2(5)

```


```{r}
#multiple arguments and their identification

pow <- function(a,p){
  a^p
}
pow(12,0.5)

```


###Excercise on function (Excercise 4)

####4.1 A quadratic equation of the generic form

\[ax^2 + bx + c =0\]

has two real roots, if they exist, given by the formula: pq formel

\[x_{1,2}=\frac{ -b \pm\sqrt{b^2 -4ac}}{2a}\]

Write a function that gives the real roots of a quadratic equation.  
Test your function on  various quadratic equations, including:
\[\begin{split}5x^2 + 6x + 1 = 0\\ 5x^2 + 2x + 1 = 0\end{split}\] 

```{r}

roots <- function (a,b,c) {
  
  if (!is.numeric( c(a,b,c)) ){
    print ("The function only makes sense with numeric values")
    
    } else if (b^2-4*a*c<0) {
  print ("no solution found")
  } else {
  
  x1 <- (-b + sqrt(b^2-4*a*c)) / 2*a
  x2 <- (-b - sqrt(b^2-4*a*c)) / 2*a
  return( c(x1, x2))
}
}

roots(a=5,c=1,b=2)
roots(a=5,b=6,c=1)
roots("tri",c=4,b=4)

```


#### 2.6-2.8 with fidalgo solution

 2.6 Create vectors with rep
For the following questions, you may want to check the help on the function rep.
Start by creating the vector
• dna = (A, C, G, T).
Then, create the following vectors:
• a = (A, C, G, T, A, C, G, T, A, C, G, T, . . . , A, C, G, T) where there are 12 A’s as shown.
• b = (A, C, G, T, A, C, G, T, A, C, G, T, . . . , A, C) where there are 15 A’s as shown; the last element
of the vector being C.
• c = (A, . . . , A, C, . . . , C, G, . . . , G, T, . . . , T) where there are 12 A’s, 15 C’s, 30 G’s and 4 T’s, in the
order shown.

```{r}

dna <- c("A","C", "G", "T")
dna

a <- rep(dna,12)


b <- rep(dna,length.out=58)
b

b2 <- rep(dna, length.out= (14*length(dna) +2))
b2

c <- rep(dna, each=c(12,15,30,4))
c
```
