---
title: "15_Solution_Poly"
author: "PGaulke"
date: "26 Juli 2019"
output: html_document
---

# Polynomials

Another very useful extension of the linear model is to include powers of variables in order to capture non-linear effects. This seems to be a contradiction in terms, but a possible answer could be that the model is still linear in the coefficients.  

To fix ideas, here is an example of fitting a quadratic model.

\[\text{mpg} = \beta_0 + \beta_{1}\times \text{horsepower} + \beta_{2}\times \text{horsepower}^2 + \varepsilon\]

This model can be estimated in the `Auto` data set of the `ISLR` package.

```{r}
require(ISLR)
data("train.data")
model.pd1 <- lm(a1 ~ NO3, data = train.data)
summary(model.pd1)
model.pd2 <- lm(a1 ~ NO3 + I(NO3^2), data = train.data)
summary(model.pd2) 
```

```{r, out.width = "100%", message = FALSE, warning = FALSE, include=TRUE, fig.cap= "Fits of mpg for various degrees of the polynomial of horsepower."}
train.data <- train.data %>%
	mutate(fit1 = predict(model.pd1),
	fit2 = predict(model.pd2),


cols <- c("Deg1", "Deg2") +
train.data %>% 
	ggplot(aes(x=a1, y=NO3)) +
	geom_point() +
	geom_line(aes(y=fit1, color="Deg.1"), size =2) +
	geom_line(aes(y=fit2, color="Deg.2"), size =2) +
	theme(legend.title = element_blank(), 
		legend.position = "bottom", 
		legend.direction = "horizontal")

```


